---
title: "Plots for Sinai Patient RNA-seq"
subtitle: "Testing response to perturbing IL10RB"
author:
- name: "[Gabriel Hoffman](http://gabrielhoffman.github.io)"
  affiliation: | 
    Icahn School of Medicine at Mount Sinai, New York
date: "Run on `r Sys.time()`"
documentclass: article
output:
  html_document:
    toc: true
    toc_float: true
params:
  upload: FALSE
---


<!---

cd /Users/gabrielhoffman/workspace/repos/sars_cov_2_npc_response/
# rm -rf analysis_cache/

rmarkdown::render("Sinai_patient_rnaseq.Rmd")


scp analysis.html sklar1:/hpc/users/hoffmg01/www/software/sars_cov_2_npc_response

--->
 
```{r setup}
library(synapser)
library(cowplot)
library(data.table)
library(ggplot2)
synLogin()
``` 

```{r read.data}
df_IL10RB = fread( synGet('syn25010953')$path )
df_IL10RB$Gene = "IL10RB"

df_IFNAR2 = fread( synGet('syn25010951')$path )
df_IFNAR2$Gene = "IFNAR2"

df = rbind(df_IL10RB, df_IFNAR2)
df$se = with(df, logFC/t) 

```{r make.plot}
df2 = df[grep("Control", df$contrast),]

lvls = c('SevereEODVsControl' = "Severe End-Organ Damage",
         'SevereVsControl' = "Severe",
        'ModerateVsControl' = "Moderate" )

df2$show = lvls[as.character(df2$contrast)]
df2$show = factor(df2$show, lvls)

ggplot(df2, aes(Gene, logFC, color=show, label=format(adj.P.Val, digits=3))) + geom_point(  position=position_dodge(width=0.9)) + theme_classic() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + coord_flip() + geom_errorbar(aes(ymin=logFC - 1.96*se, ymax=logFC + 1.96*se), width=0.1, position=position_dodge(width=0.9)) + ggtitle("Differential expression compared to controls") + geom_hline(yintercept=0, linetype="dashed") + scale_color_manual(name = "Disease state", values = c("red", "darkred", "black")) + geom_text(aes(Gene, logFC), position=position_dodge(width=0.9))
```



<!---

# 0 Control
# 1 Moderate
# 2 Severe
# 3 Severe - EODV

lvls = c(
  'ModerateVsControl' = 1 -0, 
  'SevereVsModerate'  = 2 -1, 
  'SevereVsControl'   = 2 - 0,
  'SevereEODVsSevere' = 3 - 2, 
  'SevereEODVsControl' = 3 - 0, 
  'SevereEODVsModerate' = 3-1)

lvls = sort(lvls)

df$contrast = factor(df$contrast, names(lvls))
df$distance = lvls[df$contrast]

ggplot(df, aes(Gene, logFC, color=contrast)) + geom_point(  position=position_dodge(width=0.9)) + theme_classic() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + coord_flip() + geom_errorbar(aes(ymin=logFC - 1.96*se, ymax=logFC + 1.96*se), width=0.1, position=position_dodge(width=0.9)) + ggtitle("Differential expression") + geom_hline(yintercept=0, linetype="dashed")








df2 = df[grep("Control", df$contrast),]
df2$Category = gsub("VsControl", '', as.character(df2$contrast))


df2$Expr = with(df2, AveExpr + logFC)

ggplot(df2, aes(Gene, logFC, color=Category)) + geom_point(  position=position_dodge(width=0.9)) + theme_classic() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + coord_flip() +  ggtitle("Differential expression") 





# ggplot(df, aes(contrast, logFC, fill=Gene)) + geom_bar(stat="identity", position="dodge") + theme_classic() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + coord_flip() + geom_errorbar(aes(ymin=logFC - 1.96*se, ymax=logFC + 1.96*se), width=0.1, position=position_dodge(width=0.9)) + ggtitle("Differential expression")







df_IL10RB$se = with(df_IL10RB, logFC/t) 
ggplot(df_IL10RB, aes(contrast, logFC)) + geom_bar(stat="identity") + theme_classic() + theme(aspect.ratio=1, plot.title = element_text(hjust = 0.5)) + coord_flip() + geom_errorbar(aes(ymin=logFC - 1.96*se, ymax=logFC + 1.96*se), width=0.1) + ggtitle("IL10RB")



--->












